---
title: "P8105 - Homework 1"
author: Emma Sexton <br>
date: 24 September 2022
output: github_document
---

## Problem 1
First, we need to load the packages that we will be using for this homework assignment.

```{r}
library(tidyverse)
library(palmerpenguins)
```

##### Code for Dataset Summary: 
Next, we need to load the data and briefly view the data to understand important variables and values. 

```{r}
data("penguins", package = "palmerpenguins")

summary(penguins)

head(penguins)

mean(penguins$flipper_length_mm, na.rm = TRUE)
```

##### Summary of Dataset using Inline Code:
The `penguin` dataset consists of `r nrow(penguins)` rows and `r ncol(penguins)` columns. Below, the 8 rows representing the 8 variables in the dataset are summarized using inline code. 

Variable Summary:

* *Species*: Adelie, Chinstrap, Gentoo
* *Island*: Biscoe, Dream, Torgersen
* *Bill length (mm)*: Range = `r min(penguins$bill_length_mm, na.rm = TRUE)` mm to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$bill_length_mm, na.rm = TRUE)` mm
* *Bill depth (mm)*: Range = `r min(penguins$bill_depth_mm, na.rm = TRUE)` mm to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$bill_depth_mm, na.rm = TRUE)` mm
* *Flipper length (mm)*: Range = `r min(penguins$flipper_length_mm, na.rm = TRUE)` mm to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm
* *Body mass (g)*: `r min(penguins$body_mass_g, na.rm = TRUE)` g to `r max(penguins$body_mass_g, na.rm = TRUE)` g; Mean = `r mean(penguins$body_mass_g, na.rm = TRUE)` g
* *Sex*: Female, Male
* *Year*: 2007, 2008, 2009
