---
title: "P8105 - Homework 1"
author: Emma Sexton <br>
date: 24 September 2022
output: github_document
---

## Problem 1
First, we need to load the packages that we will be using for this homework assignment.

```{r}
library(tidyverse)
library(palmerpenguins)
```

##### Code for Dataset Summary: 
Next, we need to load the data and briefly view the data to understand important variables and values. 

```{r}
data("penguins", package = "palmerpenguins")

summary(penguins)

head(penguins)

mean(penguins$flipper_length_mm, na.rm = TRUE)
```

##### Summary of Dataset using Inline Code:
The `penguin` dataset consists of `r nrow(penguins)` rows and `r ncol(penguins)` columns. Below, the 8 rows representing the 8 variables in the dataset are summarized using inline code. 

Variable Summary:

* *Species*: Adelie, Chinstrap, Gentoo
* *Island*: Biscoe, Dream, Torgersen
* *Bill length (mm)*: Range = `r min(penguins$bill_length_mm, na.rm = TRUE)` mm to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$bill_length_mm, na.rm = TRUE)` mm
* *Bill depth (mm)*: Range = `r min(penguins$bill_depth_mm, na.rm = TRUE)` mm to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$bill_depth_mm, na.rm = TRUE)` mm
* *Flipper length (mm)*: Range = `r min(penguins$flipper_length_mm, na.rm = TRUE)` mm to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm
* *Body mass (g)*: `r min(penguins$body_mass_g, na.rm = TRUE)` g to `r max(penguins$body_mass_g, na.rm = TRUE)` g; Mean = `r mean(penguins$body_mass_g, na.rm = TRUE)` g
* *Sex*: Female, Male
* *Year*: 2007, 2008, 2009

As noted above, the mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

##### Plot for `penguin` Dataset (& Save):
Next, we will create a plot of penguin flipper length vs. bill length using `ggplot` and save the plot as a PDF file. 
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point(na.rm = TRUE)

ggsave("flipper_length_vs_bill_length.pdf", w = 6, h = 4)
```

## Problem 2

##### Create a New Data Frame:
For Problem 2, we are creating a new data frame consisting of a numeric, logical, character, and factor vector using `tibble`. 
```{r}
problem_2_df <- 
  tibble(
    numeric_vector = rnorm(10),
    logic_vector = numeric_vector > 0,
    character_vector = c("Washington", "Massachusetts", "New York", "Oregon", "California", "Maine", "Vermont", "New Hampshire", "Idaho", "Colorado"),
    factor_vector = factor(c("hot", "warm", "cold", "hot", "warm", "cold", "hot", "warm", "cold", "cold"))
  )
```

##### Taking the Mean of Data Frame Variables:
Next, we will attempt to take the mean of each variable to observe what works and what does not work. 
```{r}
mean(pull(.data = problem_2_df, var = numeric_vector))
mean(pull(.data = problem_2_df, var = logic_vector))
mean(pull(.data = problem_2_df, var = character_vector))
mean(pull(.data = problem_2_df, var = factor_vector))
```

We are able to successfully generate the mean of the numeric and logic vectors; however, the character and factor vectors produced `NA` as the output.

