---
title: "P8105 - Homework 1"
author: Emma Sexton <br>
date: 24 September 2022
output: github_document
---

## Problem 1
First, we need to load the packages that we will be using for this homework assignment.

```{r}
library(tidyverse)
library(palmerpenguins)
```

##### Code for Dataset Summary: 
Next, we need to load the data and briefly view the data to understand important variables and values. 

```{r}
data("penguins", package = "palmerpenguins")

summary(penguins)

head(penguins)

mean(penguins$flipper_length_mm, na.rm = TRUE)
```

##### Summary of Dataset using Inline Code:
The `penguin` dataset consists of `r nrow(penguins)` rows and `r ncol(penguins)` columns. Below, the 8 rows representing the 8 variables in the dataset are summarized using inline code. 

Variable Summary:

* *Species*: Adelie, Chinstrap, Gentoo
* *Island*: Biscoe, Dream, Torgersen
* *Bill length (mm)*: Range = `r min(penguins$bill_length_mm, na.rm = TRUE)` mm to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$bill_length_mm, na.rm = TRUE)` mm
* *Bill depth (mm)*: Range = `r min(penguins$bill_depth_mm, na.rm = TRUE)` mm to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$bill_depth_mm, na.rm = TRUE)` mm
* *Flipper length (mm)*: Range = `r min(penguins$flipper_length_mm, na.rm = TRUE)` mm to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm; Mean = `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm
* *Body mass (g)*: `r min(penguins$body_mass_g, na.rm = TRUE)` g to `r max(penguins$body_mass_g, na.rm = TRUE)` g; Mean = `r mean(penguins$body_mass_g, na.rm = TRUE)` g
* *Sex*: Female, Male
* *Year*: 2007, 2008, 2009

As noted above, the mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

##### Plot for `penguin` Dataset (& Save):
Next, we will create a plot of penguin flipper length vs. bill length using `ggplot` and save the plot as a PDF file. 
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point(na.rm = TRUE)

ggsave("flipper_length_vs_bill_length.pdf", w = 6, h = 4)
```

## Problem 2

##### Create a New Data Frame:
For Problem 2, we are creating a new data frame consisting of a numeric, logical, character, and factor vector using `tibble`. 
```{r}
problem_2_df <- 
  tibble(
    numeric_vector = rnorm(10),
    logic_vector = numeric_vector > 0,
    character_vector = c("Washington", "Massachusetts", "New York", "Oregon", "California", "Maine", "Vermont", "New Hampshire", "Idaho", "Colorado"),
    factor_vector = factor(c("hot", "warm", "cold", "hot", "warm", "cold", "hot", "warm", "cold", "cold"))
  )
```

##### Taking the Mean of Data Frame Variables:
Next, we will attempt to take the mean of each variable to observe what works and what does not work. 
```{r}
mean(pull(.data = problem_2_df, var = numeric_vector))
mean(pull(.data = problem_2_df, var = logic_vector))
mean(pull(.data = problem_2_df, var = character_vector))
mean(pull(.data = problem_2_df, var = factor_vector))
```

We are able to successfully generate the mean of the numeric and logic vectors; however, the character and factor vectors produced `NA` as the output.

##### Using `as.numeric` to Convert Variable Types:
To test how R calculates the mean of certain variables, we will attempt to convert logical, character, and factor variables to numeric variables using `as.numeric`. 
```{r, eval=FALSE}
as.numeric(problem_2_df$logic_vector)
as.numeric(problem_2_df$character_vector)
as.numeric(problem_2_df$factor_vector)
```

The `as.numeric` function converted the `logic_vector` variable to a binary variable consisting of 0's and 1's, which explains how base R calculated the initial mean for `logic_vector`. R's decision to use 0's and 1's to represent FALSE and TRUE, respectively, explains why the mean for this variable is a decimal between 0 and 1. Since these are established values within R, it will do its best to calculate the mean for logical variables using these definitions. The `as.numeric` function does the same action that base R does when working with logical variables: applies 0's and 1's to correspond to `FALSE` and `TRUE`. 

Initially, R was unable to calculate a mean for the `factor_vector` variable because the strings 'hot', 'warm', and 'cold' did not hold numerical values from which a mean could be calculated. Using the `as_numeric` function, however, the `factor_vector` variable was converted to 1's, 2's, and 3's representing the three levels. These values were assigned alphabetically ('cold' = 1, 'hot' = 2, 'warm' = 3). Similar to the `logic_vector` variable, R needed to apply values to the `factor_vector` variable to determine the mean. It's important to be mindful of the fact that R assigns these values alphabetically and recognize that further elaboration in the code using `levels = ...` to establish the order of the levels may be required to get an accurate mean for specified levels. 

On the other hand, the `character_vector` variable was not able to be converted using the `as.numeric` function. Unlike the `factor_vector` variable, `character_vector` does not hold distinct levels or have an inherent order that R can decipher, even when converted using `as.numeric`. In summary, R is able to convert categorical ordinal variables, such as `factor_vector` to numeric variables and calculate the mean, but cannot do the same for categorical nominal variables, such as `character_vector`, thus explaining why R introduced NA by coercion, both initially and when using `as.numeric`.
